### 2022-07-04 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@146a421bb1ff8a4538f5892e0a570ea851def1bd 🚀

However, we find that the global pattern of betacoronavirus phylogenetic
distinctiveness is quite distinct from both bat host richness and
phylogenetic distinctiveness (fig.~\ref{fig:richness}; bottom). In
contrast to the sparsity of Neotropical betacoronavirus hosts, South and
Central America have the most evolutionary distinct hosts \emph{and}
viruses, followed by secondary hotspots in southeast Asia and the Rift
Valley region have mostly distinct viruses. Some degree of sampling bias
may contribute to these patterns: for example, the Neotropics are one of
the places where the fewest bat betacoronavirus sequences have been
generated (cite2), resulting in a sparser phylogenetic tree, and
artificially inflating distinctiveness; conversely, disproportionate
research effort in eastern China\textsuperscript{20} may have led to a
more complete inventory of the local diversity of coronaviruses, again
inflating these metrics relative to underlying patterns. Even accounting
for these potential biases, though, there is obvious heterogeneity in
betacoronavirus evolutionary distinctiveness that is distinct from
overall bat diversity.

Overall, these patterns recapitulate the evolutionary history of both
the order Chiroptera and the genus \emph{Betacoronavirus}. Horseshoe
bats (Rhinolophidae) include the reservoirs of the SARS-like viruses
(subgenus \emph{Sarbecovirus}), the group of pandemic threats that have
been of the greatest interest to researchers\textsuperscript{13} (and so
have been sampled most intensively).\textsuperscript{20} The hotspots of
host richness and viral diversity in southeast Asia---both of which are
disproportionately high, considering the global landscape of bat species
richness---are almost entirely driven by viral adaptive radiation
through host switching within this clade\textsuperscript{3,19}. In
contrast, the Neotropical hotspot of viral distinctiveness is driven by
isolation by host vicariance. Out of the four main groups of
betacoronaviruses, only merbecoviruses have been found in animals in the
Americas--- an introduction that is generally presumed to be
ancient.\textsuperscript{3,21} While comparatively understudied, New
World merbecoviruses have been found in the ghost-faced bats
(Mormoopidae), Neotropical leaf-nosed bats (Phyllostomidae), and
free-tailed bats (Molossidae).\textsuperscript{14--17} The former two
groups and a clade of the latter are endemic to the Neotropics, while
the explosive adaptive radiations of the phyllostomids are responsible
for the hotspot of bat diversity in the Amazon.\textsuperscript{22}
Together, these clades of New World bats play host to a distinct regime
of betacoronavirus coevolution.

@@

The existence of well-defined cophylogenetic regions suggests that the
bat-betacoronavirus system is spatially fragmented enough to create
divergent coevolutionary trajectories; in turn, this coevolutionary
mosaic may alter the risk of zoonotic emergence. These ideas are,
respectively, supported by the existence of hotspots of viral uniqueness
and the diverse origins of human betacoronaviruses. Together, this
framework points to a predictable relationship between host community
structure and coevolutionary pressure: phylogeographic structure in bat
hosts (and their diverse immune strategies)\textsuperscript{23} creates
a landscape of selective pressure; the trajectory of viruses'
coevolutionary response is, in turn, constrained by their opportunities
for either specialization or diversification through host jumps and
recombination.

@@

Compared to approaches that map emergence risk based only on the number
of known bat hosts of betacoronaviruses, our framework suggests regions
where high viral sharing dominates coevolutionary dynamics---such as
Latin America, or Eurasia above a northing of 30---would pose less of a
relative risk of zoonotic emergence. Nevertheless, areas of high host
uniqueness coupled with high viral sharing (red-to-pink in
fig.~\ref{fig:trivariate}) could create hotspots facilitated by viral
codivergence. Our framework identifies Madagascar, where most bat
species are endemic following evolutionary divergence from sister
species in both African and Asian continents,\textsuperscript{24} as one
such hotspot; interestingly, a recent study\textsuperscript{25} reported
a novel and highly divergent lineage of nobecoviruses from
Madagascar-endemic pteropid bat species (\emph{Pteropus rufus} and
\emph{Rousettus madagascariensis}), again supporting the predictive
power of the coevolutionary framework.

@@

The relationship between the underlying pathogen pool and emergence risk
is mediated by both human-wildlife interfaces (the probability of
spillover) and opportunities for onward horizontal transmission (the
probability that spillovers become epidemics)\textsuperscript{1}. As a
proxy for both, we finally overlaid the risk component from the
composite map (see above) with the proportion of built land, as a proxy
for a mix of habitat disturbance, potential for bat synanthropy or
contact with bridge hosts like livestock,\textsuperscript{26,27} and
human population density and connectivity\textsuperscript{1,28,29}
(fig.~\ref{fig:compound}). Accounting for these factors, most of South
America and Europe are at comparatively lower risk, as--although densely
populated--settlements tend to be in areas with lower potential risk.
Conversely, regions like Malaysia and the northern coast of Australia
have a high evolutionary risk component, but should represent a
relatively lower effective risk due to low human density. However,
southeast Asia, the Indian subcontinent, and scattered hotspots in
sub-Saharan Africa are at high risk due to the overlap between human
populations and natural opportunities for cross-species transmission of
betacoronaviruses.

@@

Reassuringly, these predictions correspond to the geographic origins of
the three bat-origin coronaviruses that have recently emerged in human
populations. While available information puts the spillover of
SARS-CoV-2 in a live animal market in Wuhan, China, the ultimate origin
of the virus is almost certainly in a divergent lineage of
sarbecoviruses from Indochina that was poorly characterized prior to the
pandemic.\textsuperscript{30--32} Similarly, the SARS-CoV outbreak began
in Guangdong province in 2002, reaching humans through small carnivore
bridge hosts, but was eventually traced back to a set of likely
progenitor viruses found in cave-dwelling horseshoe bats in Yunnan
province;\textsuperscript{33} nearby, antibody evidence has indicated
human exposure to SARS-like viruses.\textsuperscript{34} MERS-CoV was
first detected in Jordan, but is widespread in camels in East Africa and
the Middle East, and may have reached its bridge host decades earlier
than originally supposed;\textsuperscript{35} as a result, the geography
of the original bat-to-camel transmission is still widely regarded as
uncertain. All of these are broadly consistent with the risk factors we
identify. Notably, India and west Africa are additional hotspots that
have yet to experience the emergence of a bat coronavirus into human
populations, but may still be at risk---particularly given known gaps in
bat surveillance,\textsuperscript{20} and a dense population in both
regions with global connectivity. In any of these regions, surveillance
on viral reservoirs can be paired with targeted monitoring of high-risk
human populations (i.e., those with regular wildlife
contact)\textsuperscript{36} for maximum impact.

@@

Here, we propose a simple framework with broad explanatory power that
helps contextualize discoveries like highly divergent nobecoviruses in
Madagascar and the once-neglected adaptive radiation of sarbecoviruses
in the Indochinese peninsula. In doing so, it advances ecological theory
beyond the current state of the art for global maps of emergence risk.
For example, previous studies that have used host richness as a proxy
have predicted a high diversity of unsampled bat
viruses,\textsuperscript{18} bat coronaviruses,\textsuperscript{2} and
even specifically betacoronaviruses\textsuperscript{19} in both the
Amazon and southeast Asia. While we find that both regions are
characterized by unique and diverse communities of both hosts and
viruses, our framework is able to identify key differences between the
two systems. We find that the merbecovirus complex in Latin America has
been a unique branch of evolution separate from the rest of the global
pool, but with limited potential for viral diversification--- a finding
that is supported by previous work indicating a higher rate of
codivergence in Latin America.\textsuperscript{2} In contrast, in
southeast Asia, host richness and viral distinctiveness are high but
sharing is low; this suggests a different type of evolutionary dynamics
that could generate high local diversity of viruses through host
switching and viral recombination (see \emph{e.g.},\textsuperscript{13}
as well as the discovery of recombinant viruses with genetic material
from both the SARS-CoV and SARS-CoV-2 branches of the Sarbecovirus
lineage).\textsuperscript{45}

Both of these regions are priority areas for sampling, especially given
predictions that they contain many bat hosts of undiscovered
betacoronaviruses.\textsuperscript{19,20} However, both the evolutionary
and ecological aspects of emergence risk are higher in southeast
Asia---a fact that will only become more relevant, as bats track
shifting climates and exchange viruses with other species, creating a
hotspot of elevated cross-species transmission unique to the
region.\textsuperscript{28,46} Bats---and the spillover of their
viruses---are also sensitive to anthropogenic factors others than
climate change, including deforestation and other kinds of habitat loss,
increased stress, and greater contact with potential bridge hosts like
domesticated species.\textsuperscript{26,47--49} This represents a
challenge for both conservation strategies and pandemic
prevention,\textsuperscript{50} but identifying areas at risk, and
protecting the health of bats and ecosystems within those zones, can be
a win-win intervention for both.\textsuperscript{51--53}

@@

We downloaded the rangemap of every current bat species that was
classified as an empirically documented host of \emph{Betacoronavirus}
from the previous step, according to recent IUCN
data.\textsuperscript{54} The range maps were subsequently rasterized
using the \texttt{rasterize} function from
\texttt{GDAL}\textsuperscript{55} at a resolution of approximately
100kmx100km at the equator. For every pixel in the resulting raster
where at least one bat host of \emph{Betacoronavirus} was present, we
extract the species pool (list of all known bat hosts), which was used
to calculate the following risk assessment components: bat phylogenetic
diversity, bat compositional uniqueness, and predicted viral sharing
risk.

@@

\leavevmode\hypertarget{ref-Cohen2022SamStr}{}%
\CSLLeftMargin{20. }
\CSLRightInline{Cohen, L. E., Fagre, A. C., Chen, B., Carlson, C. J. \&
Becker, D. J. Sampling strategies and pre-pandemic surveillance gaps for
bat coronaviruses. 2022.06.15.496296 (2022)
doi:\href{https://doi.org/10.1101/2022.06.15.496296}{10.1101/2022.06.15.496296}.}

\leavevmode\hypertarget{ref-Olival2020PosRev}{}%
\CSLLeftMargin{21. }
\CSLRightInline{Olival, K. J. \emph{et al.} Possibility for reverse
zoonotic transmission of SARS-CoV-2 to free-ranging wildlife: A case
study of bats. \emph{PLoS pathogens} \textbf{16}, e1008758 (2020).}

\leavevmode\hypertarget{ref-Ammerman2012FirMol}{}%
\CSLLeftMargin{22. }
\CSLRightInline{Ammerman, L. K., Lee, D. N. \& Tipps, T. M. First
molecular phylogenetic insights into the evolution of free-tailed bats
in the subfamily Molossinae (Molossidae, Chiroptera). \emph{Journal of
Mammalogy} \textbf{93}, 12--28 (2012).}

\leavevmode\hypertarget{ref-Banerjee2020NovIns}{}%
\CSLLeftMargin{23. }
\CSLRightInline{Banerjee, A. \emph{et al.} Novel Insights Into Immune
Systems of Bats. \emph{Frontiers in Immunology} \textbf{11}, 26 (2020).}

\leavevmode\hypertarget{ref-Shi2014DeeDiv}{}%
\CSLLeftMargin{24. }
\CSLRightInline{Shi, J. J. \emph{et al.} A Deep Divergence Time between
Sister Species of Eidolon (Pteropodidae) with Evidence for Widespread
Panmixia. \emph{Acta Chiropterologica} \textbf{16}, 279--292 (2014).}

\leavevmode\hypertarget{ref-Kettenburg2022FulGen}{}%
\CSLLeftMargin{25. }
\CSLRightInline{Kettenburg, G. \emph{et al.} Full Genome Nobecovirus
Sequences From Malagasy Fruit Bats Define a Unique Evolutionary History
for This Coronavirus Clade. \emph{Frontiers in Public Health}
\textbf{10}, (2022).}

\leavevmode\hypertarget{ref-Rulli2021LanCha}{}%
\CSLLeftMargin{26. }
\CSLRightInline{Rulli, M. C., D'Odorico, P., Galli, N. \& Hayman, D. T.
Land-use change and the livestock revolution increase the risk of
zoonotic coronavirus transmission from rhinolophid bats. \emph{Nature
Food} \textbf{2}, 409--416 (2021).}

\leavevmode\hypertarget{ref-Cui2019OriEvo}{}%
\CSLLeftMargin{27. }
\CSLRightInline{Cui, J., Li, F. \& Shi, Z.-L. Origin and evolution of
pathogenic coronaviruses. \emph{Nature Reviews Microbiology}
\textbf{17}, 181--192 (2019).}

\leavevmode\hypertarget{ref-Muylaert2022PreFut}{}%
\CSLLeftMargin{28. }
\CSLRightInline{Muylaert, R. L. \emph{et al.} Present and future
distribution of bat hosts of sarbecoviruses: Implications for
conservation and public health. \emph{Proceedings of the Royal Society
B: Biological Sciences} \textbf{289}, 20220397 (2022).}

\leavevmode\hypertarget{ref-Hassell2017UrbDis}{}%
\CSLLeftMargin{29. }
\CSLRightInline{Hassell, J. M., Begon, M., Ward, M. J. \& Fèvre, E. M.
Urbanization and disease emergence: Dynamics at the
wildlife--livestock--human interface. \emph{Trends in ecology \&
evolution} \textbf{32}, 55--67 (2017).}

\leavevmode\hypertarget{ref-Worobey2022HuaMar}{}%
\CSLLeftMargin{30. }
\CSLRightInline{Worobey, M. \emph{et al.} The Huanan market was the
epicenter of SARS-CoV-2 emergence. \emph{Zenodo} (2022)
doi:\href{https://doi.org/10.5281/zenodo.6299600}{10.5281/zenodo.6299600}.}

\leavevmode\hypertarget{ref-Temmam2022BatCor}{}%
\CSLLeftMargin{31. }
\CSLRightInline{Temmam, S. \emph{et al.} Bat coronaviruses related to
SARS-CoV-2 and infectious for human cells. \emph{Nature} \textbf{604},
330--336 (2022).}

\leavevmode\hypertarget{ref-Boni2020EvoOri}{}%
\CSLLeftMargin{32. }
\CSLRightInline{Boni, M. F. \emph{et al.} Evolutionary origins of the
SARS-CoV-2 sarbecovirus lineage responsible for the COVID-19 pandemic.
\emph{Nature microbiology} \textbf{5}, 1408--1417 (2020).}

@@

\leavevmode\hypertarget{ref-Wu2021ComSur}{}%
\CSLLeftMargin{45. }
\CSLRightInline{Wu, Z. \emph{et al.} \emph{A comprehensive survey of bat
sarbecoviruses across China for the origin tracing of SARS-CoV and
SARS-CoV-2}. (2021)
doi:\href{https://doi.org/10.21203/rs.3.rs-885194/v1}{10.21203/rs.3.rs-885194/v1}.}

\leavevmode\hypertarget{ref-Carlson2022CliCha}{}%
\CSLLeftMargin{46. }
\CSLRightInline{Carlson, C. J. \emph{et al.} Climate change increases
cross-species viral transmission risk. \emph{Nature} 1--1 (2022)
doi:\href{https://doi.org/10.1038/s41586-022-04788-w}{10.1038/s41586-022-04788-w}.}

\leavevmode\hypertarget{ref-Alves2018GeoVar}{}%
\CSLLeftMargin{47. }
\CSLRightInline{Alves, D. M. C. C., Diniz-Filho, J. A. F., da Silva e
Souza, K., Gouveia, S. F. \& Villalobos, F. Geographic variation in the
relationship between large-scale environmental determinants and bat
species richness. \emph{Basic and Applied Ecology} \textbf{27}, 1--8
(2018).}

\leavevmode\hypertarget{ref-Treitler2016EffLoc}{}%
\CSLLeftMargin{48. }
\CSLRightInline{Treitler, J. T., Heim, O., Tschapka, M. \& Jung, K. The
effect of local land use and loss of forests on bats and nocturnal
insects. \emph{Ecology and Evolution} \textbf{6}, 4289--4297 (2016).}

\leavevmode\hypertarget{ref-Mendenhall2014PreBio}{}%
\CSLLeftMargin{49. }
\CSLRightInline{Mendenhall, C. D., Karp, D. S., Meyer, C. F. J., Hadly,
E. A. \& Daily, G. C. Predicting biodiversity change and averting
collapse in agricultural landscapes. \emph{Nature} \textbf{509},
213--217 (2014).}

\leavevmode\hypertarget{ref-Amman2011InvRol}{}%
\CSLLeftMargin{50. }
\CSLRightInline{Amman, B. R. \emph{et al.} \emph{Investigating the role
of bats in emerging zoonoses: Balancing ecology, conservation and public
health interest}. (FAO, 2011).}

\leavevmode\hypertarget{ref-Hopkins2021HowIde}{}%
\CSLLeftMargin{51. }
\CSLRightInline{Hopkins, S. R. \emph{et al.} How to identify win--win
interventions that benefit human health and conservation. \emph{Nature
Sustainability} \textbf{4}, 298--304 (2021).}

\leavevmode\hypertarget{ref-Plowright2021LanUse}{}%
\CSLLeftMargin{52. }
\CSLRightInline{Plowright, R. K. \emph{et al.} Land use-induced
spillover: A call to action to safeguard environmental, animal, and
human health. \emph{The Lancet Planetary Health} \textbf{5}, e237--e245
(2021).}

\leavevmode\hypertarget{ref-OHHLEP2022OneHea}{}%
\CSLLeftMargin{53. }
\CSLRightInline{OHHLEP \emph{et al.} One Health: A new definition for a
sustainable and healthy future. \emph{PLOS Pathogens} \textbf{18},
e1010537 (2022).}

\leavevmode\hypertarget{ref-IUCN2021IucRed}{}%
\CSLLeftMargin{54. }
\CSLRightInline{IUCN. The IUCN Red List of Threatened Species.
\url{https://www.iucnredlist.org} (2021).}

\leavevmode\hypertarget{ref-RouaultEven2022GdaOgr}{}%
\CSLLeftMargin{55. }
\CSLRightInline{Rouault, E. \emph{et al.} \emph{GDAL/OGR Geospatial Data
Abstraction software Library}. (Zenodo, 2022).
doi:\href{https://doi.org/10.5281/ZENODO.5884351}{10.5281/ZENODO.5884351}.}

>>>>

However, we find that the global pattern of betacoronavirus phylogenetic
distinctiveness is quite distinct from both bat host richness and
phylogenetic distinctiveness (fig.~\ref{fig:richness}; bottom). In
contrast to the sparsity of Neotropical betacoronavirus hosts, South and
Central America have the most evolutionary distinct hosts \emph{and}
viruses, followed by secondary hotspots in southeast Asia and the Rift
Valley region have mostly distinct viruses. Some degree of sampling bias
may contribute to these patterns: for example, the Neotropics are one of
the places where the fewest bat betacoronavirus sequences have been
generated,\textsuperscript{20--22} resulting in a sparser phylogenetic
tree, and artificially inflating distinctiveness; conversely,
disproportionate research effort in eastern China\textsuperscript{23}
may have led to a more complete inventory of the local diversity of
coronaviruses, again inflating these metrics relative to underlying
patterns. Even accounting for these potential biases, though, there is
obvious heterogeneity in betacoronavirus evolutionary distinctiveness
that is distinct from overall bat diversity.

Overall, these patterns recapitulate the evolutionary history of both
the order Chiroptera and the genus \emph{Betacoronavirus}. Horseshoe
bats (Rhinolophidae) include the reservoirs of the SARS-like viruses
(subgenus \emph{Sarbecovirus}), the group of pandemic threats that have
been of the greatest interest to researchers\textsuperscript{13} (and so
have been sampled most intensively).\textsuperscript{23} The hotspots of
host richness and viral diversity in southeast Asia---both of which are
disproportionately high, considering the global landscape of bat species
richness---are almost entirely driven by viral adaptive radiation
through host switching within this clade\textsuperscript{3,19}. In
contrast, the Neotropical hotspot of viral distinctiveness is driven by
isolation by host vicariance. Out of the four main groups of
betacoronaviruses, only merbecoviruses have been found in animals in the
Americas--- an introduction that is generally presumed to be
ancient.\textsuperscript{3,24} While comparatively understudied, New
World merbecoviruses have been found in the ghost-faced bats
(Mormoopidae), Neotropical leaf-nosed bats (Phyllostomidae), and
free-tailed bats (Molossidae).\textsuperscript{14--17} The former two
groups and a clade of the latter are endemic to the Neotropics, while
the explosive adaptive radiations of the phyllostomids are responsible
for the hotspot of bat diversity in the Amazon.\textsuperscript{25}
Together, these clades of New World bats play host to a distinct regime
of betacoronavirus coevolution.

@@

The existence of well-defined cophylogenetic regions suggests that the
bat-betacoronavirus system is spatially fragmented enough to create
divergent coevolutionary trajectories; in turn, this coevolutionary
mosaic may alter the risk of zoonotic emergence. These ideas are,
respectively, supported by the existence of hotspots of viral uniqueness
and the diverse origins of human betacoronaviruses. Together, this
framework points to a predictable relationship between host community
structure and coevolutionary pressure: phylogeographic structure in bat
hosts (and their diverse immune strategies)\textsuperscript{26} creates
a landscape of selective pressure; the trajectory of viruses'
coevolutionary response is, in turn, constrained by their opportunities
for either specialization or diversification through host jumps and
recombination.

@@

Compared to approaches that map emergence risk based only on the number
of known bat hosts of betacoronaviruses, our framework suggests regions
where high viral sharing dominates coevolutionary dynamics---such as
Latin America, or Eurasia above a northing of 30---would pose less of a
relative risk of zoonotic emergence. Nevertheless, areas of high host
uniqueness coupled with high viral sharing (red-to-pink in
fig.~\ref{fig:trivariate}) could create hotspots facilitated by viral
codivergence. Our framework identifies Madagascar, where most bat
species are endemic following evolutionary divergence from sister
species in both African and Asian continents,\textsuperscript{27} as one
such hotspot; interestingly, a recent study\textsuperscript{28} reported
a novel and highly divergent lineage of nobecoviruses from
Madagascar-endemic pteropid bat species (\emph{Pteropus rufus} and
\emph{Rousettus madagascariensis}), again supporting the predictive
power of the coevolutionary framework.

@@

The relationship between the underlying pathogen pool and emergence risk
is mediated by both human-wildlife interfaces (the probability of
spillover) and opportunities for onward horizontal transmission (the
probability that spillovers become epidemics)\textsuperscript{1}. As a
proxy for both, we finally overlaid the risk component from the
composite map (see above) with the proportion of built land, as a proxy
for a mix of habitat disturbance, potential for bat synanthropy or
contact with bridge hosts like livestock,\textsuperscript{29,30} and
human population density and connectivity\textsuperscript{1,31,32}
(fig.~\ref{fig:compound}). Accounting for these factors, most of South
America and Europe are at comparatively lower risk, as--although densely
populated--settlements tend to be in areas with lower potential risk.
Conversely, regions like Malaysia and the northern coast of Australia
have a high evolutionary risk component, but should represent a
relatively lower effective risk due to low human density. However,
southeast Asia, the Indian subcontinent, and scattered hotspots in
sub-Saharan Africa are at high risk due to the overlap between human
populations and natural opportunities for cross-species transmission of
betacoronaviruses.

@@

Reassuringly, these predictions correspond to the geographic origins of
the three bat-origin coronaviruses that have recently emerged in human
populations. While available information puts the spillover of
SARS-CoV-2 in a live animal market in Wuhan, China, the ultimate origin
of the virus is almost certainly in a divergent lineage of
sarbecoviruses from Indochina that was poorly characterized prior to the
pandemic.\textsuperscript{20--22} Similarly, the SARS-CoV outbreak began
in Guangdong province in 2002, reaching humans through small carnivore
bridge hosts, but was eventually traced back to a set of likely
progenitor viruses found in cave-dwelling horseshoe bats in Yunnan
province;\textsuperscript{33} nearby, antibody evidence has indicated
human exposure to SARS-like viruses.\textsuperscript{34} MERS-CoV was
first detected in Jordan, but is widespread in camels in East Africa and
the Middle East, and may have reached its bridge host decades earlier
than originally supposed;\textsuperscript{35} as a result, the geography
of the original bat-to-camel transmission is still widely regarded as
uncertain. All of these are broadly consistent with the risk factors we
identify. Notably, India and west Africa are additional hotspots that
have yet to experience the emergence of a bat coronavirus into human
populations, but may still be at risk---particularly given known gaps in
bat surveillance,\textsuperscript{23} and a dense population in both
regions with global connectivity. In any of these regions, surveillance
on viral reservoirs can be paired with targeted monitoring of high-risk
human populations (i.e., those with regular wildlife
contact)\textsuperscript{36} for maximum impact.

@@

Here, we propose a simple framework with broad explanatory power that
helps contextualize discoveries like highly divergent nobecoviruses in
Madagascar and the once-neglected adaptive radiation of sarbecoviruses
in the Indochinese peninsula. In doing so, it advances ecological theory
beyond the current state of the art for global maps of emergence risk.
For example, previous studies that have used host richness as a proxy
have predicted a high diversity of unsampled bat
viruses,\textsuperscript{18} bat coronaviruses,\textsuperscript{2} and
even specifically betacoronaviruses\textsuperscript{19} in both the
Amazon and southeast Asia. While we find that both regions are
characterized by unique and diverse communities of both hosts and
viruses, our framework is able to identify key differences between the
two systems. We find that the merbecovirus complex in Latin America has
been a unique branch of evolution separate from the rest of the global
pool, but with limited potential for viral diversification--- a finding
that is supported by previous work indicating a higher rate of
codivergence in Latin America.\textsuperscript{2,45} In contrast, in
southeast Asia, host richness and viral distinctiveness are high but
sharing is low; this suggests a different type of evolutionary dynamics
that could generate high local diversity of viruses through host
switching and viral recombination (see \emph{e.g.},\textsuperscript{13}
as well as the discovery of recombinant viruses with genetic material
from both the SARS-CoV and SARS-CoV-2 branches of the Sarbecovirus
lineage).\textsuperscript{46}

Both of these regions are priority areas for sampling, especially given
predictions that they contain many bat hosts of undiscovered
betacoronaviruses.\textsuperscript{19,23} However, both the evolutionary
and ecological aspects of emergence risk are higher in southeast
Asia---a fact that will only become more relevant, as bats track
shifting climates and exchange viruses with other species, creating a
hotspot of elevated cross-species transmission unique to the
region.\textsuperscript{31,47} Bats---and the spillover of their
viruses---are also sensitive to anthropogenic factors others than
climate change, including deforestation and other kinds of habitat loss,
increased stress, and greater contact with potential bridge hosts like
domesticated species.\textsuperscript{29,48--50} This represents a
challenge for both conservation strategies and pandemic
prevention,\textsuperscript{51} but identifying areas at risk, and
protecting the health of bats and ecosystems within those zones, can be
a win-win intervention for both.\textsuperscript{52--54}

@@

We downloaded the rangemap of every current bat species that was
classified as an empirically documented host of \emph{Betacoronavirus}
from the previous step, according to recent IUCN
data.\textsuperscript{55} The range maps were subsequently rasterized
using the \texttt{rasterize} function from
\texttt{GDAL}\textsuperscript{56} at a resolution of approximately
100kmx100km at the equator. For every pixel in the resulting raster
where at least one bat host of \emph{Betacoronavirus} was present, we
extract the species pool (list of all known bat hosts), which was used
to calculate the following risk assessment components: bat phylogenetic
diversity, bat compositional uniqueness, and predicted viral sharing
risk.

@@

\leavevmode\hypertarget{ref-Worobey2022HuaMar}{}%
\CSLLeftMargin{20. }
\CSLRightInline{Worobey, M. \emph{et al.} The Huanan market was the
epicenter of SARS-CoV-2 emergence. \emph{Zenodo} (2022)
doi:\href{https://doi.org/10.5281/zenodo.6299600}{10.5281/zenodo.6299600}.}

\leavevmode\hypertarget{ref-Temmam2022BatCor}{}%
\CSLLeftMargin{21. }
\CSLRightInline{Temmam, S. \emph{et al.} Bat coronaviruses related to
SARS-CoV-2 and infectious for human cells. \emph{Nature} \textbf{604},
330--336 (2022).}

\leavevmode\hypertarget{ref-Boni2020EvoOri}{}%
\CSLLeftMargin{22. }
\CSLRightInline{Boni, M. F. \emph{et al.} Evolutionary origins of the
SARS-CoV-2 sarbecovirus lineage responsible for the COVID-19 pandemic.
\emph{Nature microbiology} \textbf{5}, 1408--1417 (2020).}

\leavevmode\hypertarget{ref-Cohen2022SamStr}{}%
\CSLLeftMargin{23. }
\CSLRightInline{Cohen, L. E., Fagre, A. C., Chen, B., Carlson, C. J. \&
Becker, D. J. Sampling strategies and pre-pandemic surveillance gaps for
bat coronaviruses. 2022.06.15.496296 (2022)
doi:\href{https://doi.org/10.1101/2022.06.15.496296}{10.1101/2022.06.15.496296}.}

\leavevmode\hypertarget{ref-Olival2020PosRev}{}%
\CSLLeftMargin{24. }
\CSLRightInline{Olival, K. J. \emph{et al.} Possibility for reverse
zoonotic transmission of SARS-CoV-2 to free-ranging wildlife: A case
study of bats. \emph{PLoS pathogens} \textbf{16}, e1008758 (2020).}

\leavevmode\hypertarget{ref-Ammerman2012FirMol}{}%
\CSLLeftMargin{25. }
\CSLRightInline{Ammerman, L. K., Lee, D. N. \& Tipps, T. M. First
molecular phylogenetic insights into the evolution of free-tailed bats
in the subfamily Molossinae (Molossidae, Chiroptera). \emph{Journal of
Mammalogy} \textbf{93}, 12--28 (2012).}

\leavevmode\hypertarget{ref-Banerjee2020NovIns}{}%
\CSLLeftMargin{26. }
\CSLRightInline{Banerjee, A. \emph{et al.} Novel Insights Into Immune
Systems of Bats. \emph{Frontiers in Immunology} \textbf{11}, 26 (2020).}

\leavevmode\hypertarget{ref-Shi2014DeeDiv}{}%
\CSLLeftMargin{27. }
\CSLRightInline{Shi, J. J. \emph{et al.} A Deep Divergence Time between
Sister Species of Eidolon (Pteropodidae) with Evidence for Widespread
Panmixia. \emph{Acta Chiropterologica} \textbf{16}, 279--292 (2014).}

\leavevmode\hypertarget{ref-Kettenburg2022FulGen}{}%
\CSLLeftMargin{28. }
\CSLRightInline{Kettenburg, G. \emph{et al.} Full Genome Nobecovirus
Sequences From Malagasy Fruit Bats Define a Unique Evolutionary History
for This Coronavirus Clade. \emph{Frontiers in Public Health}
\textbf{10}, (2022).}

\leavevmode\hypertarget{ref-Rulli2021LanCha}{}%
\CSLLeftMargin{29. }
\CSLRightInline{Rulli, M. C., D'Odorico, P., Galli, N. \& Hayman, D. T.
Land-use change and the livestock revolution increase the risk of
zoonotic coronavirus transmission from rhinolophid bats. \emph{Nature
Food} \textbf{2}, 409--416 (2021).}

\leavevmode\hypertarget{ref-Cui2019OriEvo}{}%
\CSLLeftMargin{30. }
\CSLRightInline{Cui, J., Li, F. \& Shi, Z.-L. Origin and evolution of
pathogenic coronaviruses. \emph{Nature Reviews Microbiology}
\textbf{17}, 181--192 (2019).}

\leavevmode\hypertarget{ref-Muylaert2022PreFut}{}%
\CSLLeftMargin{31. }
\CSLRightInline{Muylaert, R. L. \emph{et al.} Present and future
distribution of bat hosts of sarbecoviruses: Implications for
conservation and public health. \emph{Proceedings of the Royal Society
B: Biological Sciences} \textbf{289}, 20220397 (2022).}

\leavevmode\hypertarget{ref-Hassell2017UrbDis}{}%
\CSLLeftMargin{32. }
\CSLRightInline{Hassell, J. M., Begon, M., Ward, M. J. \& Fèvre, E. M.
Urbanization and disease emergence: Dynamics at the
wildlife--livestock--human interface. \emph{Trends in ecology \&
evolution} \textbf{32}, 55--67 (2017).}

@@

\leavevmode\hypertarget{ref-Caraballo2022CroTra}{}%
\CSLLeftMargin{45. }
\CSLRightInline{Caraballo, D. A. Cross-Species Transmission of Bat
Coronaviruses in the Americas: Contrasting Patterns between
Alphacoronavirus and Betacoronavirus. \emph{Microbiology Spectrum}
\textbf{0}, e01411--22 (2022).}

\leavevmode\hypertarget{ref-Wu2021ComSur}{}%
\CSLLeftMargin{46. }
\CSLRightInline{Wu, Z. \emph{et al.} \emph{A comprehensive survey of bat
sarbecoviruses across China for the origin tracing of SARS-CoV and
SARS-CoV-2}. (2021)
doi:\href{https://doi.org/10.21203/rs.3.rs-885194/v1}{10.21203/rs.3.rs-885194/v1}.}

\leavevmode\hypertarget{ref-Carlson2022CliCha}{}%
\CSLLeftMargin{47. }
\CSLRightInline{Carlson, C. J. \emph{et al.} Climate change increases
cross-species viral transmission risk. \emph{Nature} 1--1 (2022)
doi:\href{https://doi.org/10.1038/s41586-022-04788-w}{10.1038/s41586-022-04788-w}.}

\leavevmode\hypertarget{ref-Alves2018GeoVar}{}%
\CSLLeftMargin{48. }
\CSLRightInline{Alves, D. M. C. C., Diniz-Filho, J. A. F., da Silva e
Souza, K., Gouveia, S. F. \& Villalobos, F. Geographic variation in the
relationship between large-scale environmental determinants and bat
species richness. \emph{Basic and Applied Ecology} \textbf{27}, 1--8
(2018).}

\leavevmode\hypertarget{ref-Treitler2016EffLoc}{}%
\CSLLeftMargin{49. }
\CSLRightInline{Treitler, J. T., Heim, O., Tschapka, M. \& Jung, K. The
effect of local land use and loss of forests on bats and nocturnal
insects. \emph{Ecology and Evolution} \textbf{6}, 4289--4297 (2016).}

\leavevmode\hypertarget{ref-Mendenhall2014PreBio}{}%
\CSLLeftMargin{50. }
\CSLRightInline{Mendenhall, C. D., Karp, D. S., Meyer, C. F. J., Hadly,
E. A. \& Daily, G. C. Predicting biodiversity change and averting
collapse in agricultural landscapes. \emph{Nature} \textbf{509},
213--217 (2014).}

\leavevmode\hypertarget{ref-Amman2011InvRol}{}%
\CSLLeftMargin{51. }
\CSLRightInline{Amman, B. R. \emph{et al.} \emph{Investigating the role
of bats in emerging zoonoses: Balancing ecology, conservation and public
health interest}. (FAO, 2011).}

\leavevmode\hypertarget{ref-Hopkins2021HowIde}{}%
\CSLLeftMargin{52. }
\CSLRightInline{Hopkins, S. R. \emph{et al.} How to identify win--win
interventions that benefit human health and conservation. \emph{Nature
Sustainability} \textbf{4}, 298--304 (2021).}

\leavevmode\hypertarget{ref-Plowright2021LanUse}{}%
\CSLLeftMargin{53. }
\CSLRightInline{Plowright, R. K. \emph{et al.} Land use-induced
spillover: A call to action to safeguard environmental, animal, and
human health. \emph{The Lancet Planetary Health} \textbf{5}, e237--e245
(2021).}

\leavevmode\hypertarget{ref-OHHLEP2022OneHea}{}%
\CSLLeftMargin{54. }
\CSLRightInline{OHHLEP \emph{et al.} One Health: A new definition for a
sustainable and healthy future. \emph{PLOS Pathogens} \textbf{18},
e1010537 (2022).}

\leavevmode\hypertarget{ref-IUCN2021IucRed}{}%
\CSLLeftMargin{55. }
\CSLRightInline{IUCN. The IUCN Red List of Threatened Species.
\url{https://www.iucnredlist.org} (2021).}
